---
title: "Simple Sorption Kinetic model"
output: html_document
---
# Logistic Setup

Test text

## Package dependencies

```{r}
library(deSolve)
library("FME")
library(dplyr)
library(tidyr)
library(ggplot2)

```

## Set Working Directory

```{r}
setwd("C:/Users/Hannah/Desktop/kinetics/Simple Model")

```


# Model structure

## Parameter names & Equations 

### Change in mass over time (mg/hr)
- dDS = change in Dismal Swamp DOC mass in solution over time
- dLT = change in wetland CDOC mass in solution over time
- dNC = change in wetland NCDOC mass in solution over time
### Rate parameters (hr-1)
- kdDS, kdLT, kdNC = desorption rate
- kaDS, kaLT, kaNC = adsorption rate
### State Variables (mg)
- DSs, LTs, NCs = sediment Dismal Swamp DOC, wetland CDOC, and wetland NCDOC mass
- DSw, LTw, NCw = water Dismal Swamp DOC, wetland CDOC, and wetland NCDOC mass
### First order ODEs 
- dDSw <- kdDS*DSs - kaDS*DSw #del conc of GDS DOC in solution
- dLTw <- kdLT*LTs - kaLT*LTw #del conc of sediment CDOC in solution -- colored
- dNCw <- kdNC*NCs - kaNC*NCw #del conc of sediment NCDOC in solution -- non-colored
- dDSs <- kaDS*DSw - kdDS*DSs #del conc of GDS DOC on sediment
- dLTs <- kaLT*LTw - kdLT*LTs #del conc of sediment CDOC on sediment -- colored 
- dNCs <- kaNC*NCw - kdNC*NCs #del conc of sediment NCDOC on sediment -- non-colored
- change in mass over time is equal to the addition from desorption and loss via adsorption at every time step
### ID names
- JB__ or TA__ = kinetic experiments
  - JB = Jug Bay
  - TA = Taskinas
- W___ or P___ = spatial kinetic experiments
  - W = shallow, P = deep
  - C = creek edge, I = intermediate site, M = high marsh
- __HF, __HS, __LF, __LS = scenario
  - HF = High initial [DOC], Fresh
  - HS = High initial [DOC], Saline
  - LF = Low initial [DOC], Fresh
  - LS = Low initial [DOC], Saline
- 1-7 = time points
  - 1 = 0 min, 2 = 10 min, 3 = 15 min, 4 = 1 hr, 5 = 6 hr, 6 = 12 hr, 7 = 24 hr


## Compartment Model 
### List the six state variable ODEs
```{r}
threecomp <- function (time, y, parms, ...) {
   with(as.list(c(parms, y)), {
   dDSw <- kdDS*DSs - kaDS*DSw 
   dLTw <- kdLT*LTs - kaLT*LTw 
   dNCw <- kdNC*NCs - kaNC*NCw 
   dDSs <- kaDS*DSw - kdDS*DSs 
   dLTs <- kaLT*LTw - kdLT*LTs 
   dNCs <- kaNC*NCw - kdNC*NCs 
   list(c(dDSw, dLTw, dNCw, dDSs, dLTs, dNCs))})}
```

# Data organization

## Time

- Time will always stay the same 7 values: 0, 0.17, 0.25, 1, 6, 12, 24
- hours

## Observed values

- Only have measured observed values for DSw, LTw, and NCw 
- These will change with every scenario -- need to input these data

## Initial Values for model

- Six y0 values for DSw, LTw, NCw, DSs, LTs, NCs
- These will change with every scenario -- need to input these data
- Rate parameter k's might change but the first guess could be the same for each

### Using dplyr to manipulate the data
- Try to use .csv as default file formats
- Try not to have spaces or unconventional characters in title names
- Remember that initial values must be in  the same order as the ODE order (Use %>% select())
- data.frame <- as.numeric changes data to vector
- Use str() to compare the data types 

## Read in time, y0, and obs values
```{r}

time_pts <- c(0, 0.17, 0.25, 1, 6, 12, 24)

y0df <- read.csv("~/Downloads/y0_master.csv")
JHF <- y0df %>% filter(X == "JBHF")
str(y0df)
#vector <- as.numeric(y0df) 

obs <- read.csv("~/Downloads/simpmodel_obs.csv")


```

## Use dplyr and tidyr to select the observational data for the experiment
```{r}
obs2 <- obs %>% separate(X, c("exp", "time"), 4)
JHF <- obs2 %>% filter(exp == "JBHF")
# obs3 <- obs2 %>% mutate(time = c(0,0.17,0.25,1,6,12,24)) -- adds time to data frame if it wasn't already there
```

# Rate Parameters
```{r}
parms <- c(kdDS = 0.74, kaDS = 0.35, kdLT = -3.9e-13, kaLT = 34.9, kdNC = 13.2,  kaNC = -11.5) #guess at k's
times <- seq(0, 24, length=200)
y0 <- c(DSw = 8.867, LTw = 0, NCw = 0, DSs = 0, LTs = 0, NCs = 0.16)
out1 <- ode(y0, times, threecomp, parms)
plot(out1) #), obs=dat)
```

# Cost Function 
```{r}
cost <- function(p, dat) {
    out <- ode(y0, times, threecomp, p)
    modCost(out, dat, weight = "none")} # try weight = "std" or "mean"
```
 
# Fit the Model 
```{r}
fit <- modFit(f = cost, p = parms)
 summary(fit)
```

```{r}



#ans <- lapply(1:nrow(y0df), function(x){
ans <- lapply(c(1,3), function(x){
  site <- as.character(y0df$X[x])
  dat <- obs2 %>% filter(exp == site) %>% mutate(time = time_pts)
  print(dat)
  parms <- c(kdDS = 0.74, kaDS = 0.35, kdLT = -3.9e-13, kaLT = 34.9, kdNC = 13.2,  kaNC = -11.5) #guess at k's
  times <- seq(0, 24, length=200)
  y0 <- c(DSw = 8.867, LTw = 0, NCw = 0, DSs = 0, LTs = 0, NCs = 0.16) # filter out from y0 master
  # make sure when filter as data.frame to convert from data.frame to named numeric
  # use str to check
  out1 <- ode(y0, times, threecomp, parms)
  plot(out1) #), obs=dat)
  dat2 = dat %>% select(time, DSw, LTw, NCw)
  fit <- modFit(f = cost, p = parms, dat = dat2)
  summary(fit)
  par = data.frame(as.list(fit$par)) %>% mutate(X = site)
  return(par)
})

all_ans <- bind_rows(ans) #binds all the rows into one data.frame that can be saved to csv
# check what this means: Cannot estimate covariance; system is singular
# try starting with different intial parameter values to see if there are different local minima for best parameters
# may want to make our own plots next week - yaay ggplot!
```

# Jacqueline should show Hannah lapply
```{r}
# x <- lapply(1:nrow(df), function(x){
#   print(x)
#   # 1. pull in observed data, separate by scenario
#   # 2. pull in y0 data, separate by scenario
#   # 3. determine rate parameters
#   # 4. ODE function
#   # 5. plot output of six state variables
#   # 6. plot output of three state variables with obs data
#   # 7. save plots
#   # 8. cost function
#   # 9. model fit
#   # 10. calc, plot residuals for three state variables, save plot
#   df[x,]
# })

rates <- sapply(y0$X, function(rates){
  exp <- y0 %>% filter(X == "X")
  obsv <- obs2 %>% filter(exp == "X")
  parms <- c(kdDS = 0.74, kaDS = 0.35, kdLT = -3.9e-13, kaLT = 34.9, kdNC = 13.2,  kaNC = -11.5) #guess at k's
  times <- seq(0, 24, length=200)
  out1 <- ode(y0, times, threecomp, parms)
  obsplot <- plot(out1, obs=dat)
  cost <- function(p) {
  out <- ode(y0, times, threecomp, p)
  modCost(out, dat, weight = "none")}
  fit <- modFit(f = cost, p = parms)
  sum <- summary(fit)
})

x
```


# Plots
## plot model & observations for DSw, LTw, NCw, TDOC concentrations over time
## solve for and plot residuals for the above to find patters
## save each, move on to next scenario

## 32 total scenarios -- data sheets have 8 as examples