---
title: "Simple Sorption Kinetic model"
output: html_document
---
# Logistic Setup

## Package dependencies
setwd("~/Desktop/Professional/sedsorption_model")
```{r}
devtools::load_all()
library(deSolve)
library("FME")
library(dplyr)
library(tidyr)
library(ggplot2)

```


# Model structure

## Parameter names & Equations 

### Change in mass over time (mg/hr)
- dDS = change in Dismal Swamp DOC mass in solution over time
- dLT = change in wetland CDOC mass in solution over time
- dNC = change in wetland NCDOC mass in solution over time
### Rate parameters (hr-1)
- kdDS, kdLT, kdNC = desorption rate
- kaDS, kaLT, kaNC = adsorption rate
### State Variables (mg)
- DSs, LTs, NCs = sediment Dismal Swamp DOC, wetland CDOC, and wetland NCDOC mass
- DSw, LTw, NCw = water Dismal Swamp DOC, wetland CDOC, and wetland NCDOC mass
### First order ODEs 
- dDSw <- kdDS * DSs - kaDS * DSw #del mass of GDS DOC in solution
- dLTw <- kdLT * LTs - kaLT * LTw #del mass of sediment CDOC in solution -- colored
- dNCw <- kdNC * NCs - kaNC * NCw #del mass of sediment NCDOC in solution -- non-colored
- dDSs <- kaDS * DSw - kdDS * DSs #del mass of GDS DOC on sediment
- dLTs <- kaLT * LTw - kdLT * LTs #del mass of sediment CDOC on sediment -- colored 
- dNCs <- kaNC * NCw - kdNC * NCs #del mass of sediment NCDOC on sediment -- non-colored
- change in mass over time is equal to the addition from desorption and loss via adsorption at every time step
### ID names
- JB__ or TA__ = kinetic experiments
  - JB = Jug Bay
  - TA = Taskinas
- W___ or P___ = spatial kinetic experiments
  - W = shallow, P = deep
  - C = creek edge, I = intermediate site, M = high marsh
- __HF, __HS, __LF, __LS = scenario
  - HF = High initial [DOC], Fresh
  - HS = High initial [DOC], Saline
  - LF = Low initial [DOC], Fresh
  - LS = Low initial [DOC], Saline
- 1-7 = time points
  - 1 = 0 min, 2 = 10 min, 3 = 15 min, 4 = 1 hr, 5 = 6 hr, 6 = 12 hr, 7 = 24 hr


## Compartment Model 
### List the six state variable ODEs
```{r}
threecomp <- function (time, y, parms, ...) {
   with(as.list(c(parms, y)), {
   dDSw <- kdDS*DSs - kaDS*DSw 
   dLTw <- kdLT*LTs - kaLT*LTw 
   dNCw <- kdNC*NCs - kaNC*NCw 
   dDSs <- kaDS*DSw - kdDS*DSs 
   dLTs <- kaLT*LTw - kdLT*LTs 
   dNCs <- kaNC*NCw - kdNC*NCs
   list(c(dDSw, dLTw, dNCw, dDSs, dLTs, dNCs))})}

twocomp <- function (time, y, parms, ...) {
   with(as.list(c(parms, y)), {
   dDSw <- kdDS*DSs - kaDS*DSw 
   dNCw <- kdNC*NCs - kaNC*NCw 
   dDSs <- kaDS*DSw - kdDS*DSs 
   dNCs <- kaNC*NCw - kdNC*NCs
   list(c(dDSw, dNCw, dDSs, dNCs))})}
```

# Data organization
## Time

- Time will always stay the same 7 values: 0, 0.17, 0.25, 1, 6, 12, 24
- hours

## Observed values

- Only have measured observed values for DSw, LTw, and NCw 
- These will change with every scenario -- need to input these data

## Initial Values for model

- Six y0 values for DSw, LTw, NCw, DSs, LTs, NCs
- These will change with every scenario -- need to input these data
- Rate parameter k's might change but the first guess could be the same for each

### Using dplyr to manipulate the data
- Try to use .csv as default file formats
- Try not to have spaces or unconventional characters in title names
- Remember that initial values must be in  the same order as the ODE order (Use %>% select())
- data.frame <- as.numeric changes data to vector
- Use str() to compare the data types 

## Read in time, y0, and obs values
```{r}

time_pts <- c(0, 0.17, 0.25, 1, 6, 12, 24)

y0df <- read.csv("C:/Users/Hannah/Desktop/kinetics/Simple Model/y0_master.csv")
y0df <- y0df[,c(1,2,4,5,7)]

obs <- read.csv("C:/Users/Hannah/Desktop/kinetics/Simple Model/simpmodel_obs.csv")

obs2 <- obs %>% separate(X, c("exp", "time"), 4)


```

## Use dplyr and tidyr to select the observational data for the experiment
### Examples

- #JHF <- y0df %>% filter(X == "JBHF")
- obs2 <- obs %>% separate(X, c("exp", "time"), 4) -- breaks one column into two at the fourth character
- #JHF <- obs2 %>% filter(exp == "JBHF")
- # obs3 <- obs2 %>% mutate(time = c(0,0.17,0.25,1,6,12,24)) -- adds time to data frame if it wasn't already there

# Rate Parameters
```{r}
# The sequence for the model - needs the rates, model time output, y0 values

#parms <- c(kdDS = 0.74, kaDS = 0.35, kdLT = -3.9e-13, kaLT = 34.9, kdNC = 13.2,  kaNC = -11.5) #guess at k's
#times <- seq(0, 24, length=200)
#y0 <- c(DSw = 8.867, LTw = 0, NCw = 0, DSs = 0, LTs = 0, NCs = 0.16)
#out1 <- ode(y0, times, threecomp, parms)
#plot(out1) #), obs=dat)
```

# Cost Function 
```{r}
cost <- function(p, dat) {
    out <- ode(y0, times, threecomp, p)
    modCost(out, dat, weight = "none")} # try weight = "std" or "mean"
```
 
# Fit the Model 
```{r}
#fit <- modFit(f = cost, p = parms)
 #summary(fit)
```

```{r}

#ans <- lapply(1:nrow(y0df), function(x){    #do all rows
#ans <- lapply(c(1,3), function(x){          #try just rows 1 and 3
site = "JBHF"                                #try just one site 
#  site <- as.character(y0df$X[x])
  dat <- obs2 %>% filter(exp == site) %>% mutate(time = time_pts)
  print(dat)
  #parms <- c(kdDS = 0.1, kaDS = 0.1, kdLT = 0, kaLT = 0, kdNC = 0.1,  kaNC = 0.1) #guess at k's
  parms <- c(kdDS = 0.1, kaDS = 0.1, kdNC = 0.1,  kaNC = 0.1) #guess at k's
  print(parms)
  times <- seq(0, 24, length=200)
  y0 <- y0df %>% filter(X == site) # filter out from y0 master
  y0 <- data.matrix(y0) #makes the data frame a numeric
  y0 <- y0[,-1] #get rid of X label column
  print(y0)
  dat2 = dat %>% select(time, DSw, LTw, NCw)
 
  dat3 <- dat2 %>% select_if(~sum(.)!=0)
  #takes dat2 and looks for a column whose sum is 0 then keeps columns that are not
  
  out1 <- ode(y0, times, threecomp, parms)
  plot(out1) #), obs=dat)
  fit <- modFit(f = cost, p = parms, dat = dat3)
  summary(fit)
  
  par = data.frame(as.list(fit$par)) %>% mutate(X = site)
  #return(par)
#})

all_ans <- bind_rows(ans) #binds all the rows into one data.frame that can be saved to csv
# check what this means: Cannot estimate covariance; system is singular
# try starting with different initial parameter values to see if there are different local minima for best parameters
# may want to make our own plots next week - yaay ggplot!
```

```{r}
#rm(list = ls())  # Clean up

```


# Plots
## plot model & observations for DSw, LTw, NCw, TDOC concentrations over time
## solve for and plot residuals for the above to find patters
## save each, move on to next scenario

## 32 total scenarios -- data sheets have 8 as examples